---
title: "analysis"
---
First we load the libraries needed for the analysis: 
```{r}
#| output: false
library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)
```
Begin looking at the data. The first 10 lines of the babynames dataset looks like: 
```{r}
head(babynames) |> kable()
```
Create the functions to: 
* Get the most frequent babynames over a given time period
* From this output, plot the top n most popular names

```{r}
#| code-fold: true
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

Plot the girl names: 
```{r}
#| label: "fig-top-girl-names"
#| fig-cap: "Top girl babynames over the years"
#| fig-align: "left"
get_most_frequent(babynames, select_sex = "F") |>
  plot_top()
```
From @fig-top-girl-names you can see that the most popular name around the 1980s was Jennifer, followed by Lisa, Jessica and Ashley. 

Plot the boy names: 
```{r}
#| label: "fig-top-boy-names"
#| fig-cap: "Top boy babynames over the years"
#| fig-align: "left"
get_most_frequent(babynames, select_sex = "M") |>
  plot_top()
```
From @fig-top-boy-names you can see that the most popular name around the 1980s was Michael. 
